Estimating duration in the presence of misclassification

### 2024-03-21 biology-data.tex
add SRS details

I modified the regression component of a pre-existing version of a MRP model for CIS~\autocite{pouwelsMRPvaccination}, as recommended by that model's author for the specific context of this thesis~\citePersonalComms{Koen Pouwels}.
As in prior work~\autocite{pouwelsMRPvaccination}, I used INLA for inference, implemented in R-INLA (see \cref{E-transmission:sec:INLA}).
For a day $t$, age group $a$, ethnicity $e$, and sex $s$ the regression model is:
\begin{align}
    \logit(p_{taes}) &= \beta + \beta_s + \beta_{w(e)} + u_a + v_e + w_t + l_{a,t} \\
    y_{taes} &\dist \BB(n_{taes}, p_{taes}, \rho).
\end{align}
The terms are as follows.
\todo[inline]{Check I've written the priors correctly}
\begin{itemize}
    \item $n_{taes}$ and $y_{taes}$ are the number of individuals tested and the number of positive test results on day $t$ in individuals of age group $a$, ethnicity $e$, and sex $s$. I use a beta-binomial distribution, as discussed in \cref{biology-data:sec:clustering}.
    \item $\beta$ is a global intercept term with the vague prior $\beta \dist \N(0, 1/0.001^2)$.
    \item $\beta_s$ is a fixed effect for sex, with $\beta_s = 0$ for males and estimated for females. The prior for $\beta_s$ is vague, $\beta_s \dist \N(0, 1/0.001^2)$.
    \item $\beta_{w(e)}$ is a fixed effect for White or non-White ethnicity. $\beta_{w(e)} = 0$ if $e$ corresponds to White ethnicity and $\beta_{w(e)} = \beta_{n}$ otherwise (\ie non-White ethnic groups).
      The prior for $\beta_n$ is vague, $\beta_n \dist \N(0, 1/0.001^2)$.
      % $\beta_n$ and $\beta_f$ have the default INLA prior for fixed effects\todo{check what this is}.
    \item $u_a$ and $v_e$ are independent normal random effects for age and ethnic group respectively. Their standard deviation has a penalized complexity prior~\autocite{simpsonPenalising} with a 10\% probability of being greater than 1.
    \item $w_t$ is a second-order random walk, that is $w_t = 2w_{t-1} - w_{t-2} + \epsilon_t$ where $\epsilon_t \dist \N(0, \sigma_w^2)$.
      $\sigma_w$ has a vague prior, $1/\sigma_w \dist \GamDist(1, 5 \times 10^{-5})$.
    \item $l_{a,t}$ is an age-group specific second-order random walk.
      That is $l_{a,t} = 2l_{a,t-1} - l_{a,t-2} + \epsilon_{a,t}$ where $\epsilon_{a,t} \dist \N(0, \sigma_l^2)$.
      $\sigma_l$ has a vague prior, $1/\sigma_l \dist \GamDist(1, 5 \times 10^{-5})$.
\end{itemize}
For parameters given vague priors, the large amount of data means that the prior has little influence on the posterior distribution.
The second-order random walk priors favours linear growth in $\logit(p_{taes})$ over time.
For the small values of $p_{taes}$ likely to be encountered, this approximates exponential growth.

I fitted the model separately and independently for each region, implicitly adding all interactions between region and the other variables.
This step assumes that the estimates for each region can be considered independent, which is reasonable because the regions are large with many tests conducted within each.

The final step is poststratifying the incidence proportion to the CIS target population and subpopulations of interest.
To start, I draw a sample of 500 prevalence time series in each stratum from the approximate posterior distribution (using the procedure described in \cref{transmission:sec:INLA:posterior}).
For each posterior draw and day, the incidence proportion in the (sub)population is the weighted average of the incidence proportion in each stratum, where the weights are the number of individuals in that stratum.
The population numbers for each stratum are provided by the Office for National Statistics.

@@

Analyses requiring more fine-grained stratification (such as that in \cref{biology-data:sec:MRP}) or linking data from the same individual over time (such as that in \cref{E-perf-test,E-imperf-test}) must be performed within the SRS.
Therefore, computationally efficient approaches must be taken to this work.
The SRS contains limited computational power.
It has\todo{insert SRS computational power}.
However, these are shared computational resources and due consideration must be given to other users.
Processes which run for multiple days are liable to be cancelled.

This computational limitation will feature in several places in this thesis.

>>>>

I modified the regression component of a pre-existing version of a MRP model for CIS~\autocite{pouwelsMRPvaccination}, as recommended by that model's author for the specific context of this thesis~\citePersonalComms{Koen Pouwels}.
As in prior work~\autocite{pouwelsMRPvaccination}, I used INLA for inference, implemented in R-INLA (see \cref{E-transmission:sec:INLA}).
For a day $t$, age group $a$, ethnicity $e$, and sex $s$ the regression model is:
\begin{align}
    \logit(p_{taes}) &= \beta + \beta_s + \beta_{w(e)} + u_a + v_e + w_t + l_{a,t} \\
    y_{taes} &\dist \BB(n_{taes}, p_{taes}, \rho).
\end{align}
The terms are as follows.
\todo[inline]{Check I've written the priors correctly}
\begin{itemize}
    \item $n_{taes}$ and $y_{taes}$ are the number of individuals tested and the number of positive test results on day $t$ in individuals of age group $a$, ethnicity $e$, and sex $s$. I use a beta-binomial distribution for the likelihood, as discussed in \cref{biology-data:sec:clustering}.
    \item $\beta$ is a global intercept term with the vague prior $\beta \dist \N(0, 1/0.001^2)$.
    \item $\beta_s$ is a fixed effect for sex, with $\beta_s = 0$ for males and estimated for females. The prior for $\beta_s$ is vague, $\beta_s \dist \N(0, 1/0.001^2)$.
    \item $\beta_{w(e)}$ is a fixed effect for White or non-White ethnicity. $\beta_{w(e)} = 0$ if $e$ corresponds to White ethnicity and $\beta_{w(e)} = \beta_{n}$ otherwise (\ie non-White ethnic groups).
      The prior for $\beta_n$ is vague, $\beta_n \dist \N(0, 1/0.001^2)$.
      % $\beta_n$ and $\beta_f$ have the default INLA prior for fixed effects\todo{check what this is}.
    \item $u_a$ and $v_e$ are independent normal random effects for age and ethnic group respectively. Their standard deviation has a penalized complexity prior~\autocite{simpsonPenalising} with a 10\% probability of being greater than 1.
    \item $w_t$ is a second-order random walk, that is $w_t = 2w_{t-1} - w_{t-2} + \epsilon_t$ where $\epsilon_t \dist \N(0, \sigma_w^2)$.
      $\sigma_w$ has a vague prior, $1/\sigma_w \dist \GamDist(1, 5 \times 10^{-5})$.
    \item $l_{a,t}$ is an age-group specific second-order random walk.
      That is $l_{a,t} = 2l_{a,t-1} - l_{a,t-2} + \epsilon_{a,t}$ where $\epsilon_{a,t} \dist \N(0, \sigma_l^2)$.
      $\sigma_l$ has a vague prior, $1/\sigma_l \dist \GamDist(1, 5 \times 10^{-5})$.
\end{itemize}
% For parameters given vague priors, the large amount of data means that the prior has little influence on the posterior distribution.
The second-order random walk priors favours linear growth in $\logit(p_{taes})$ over time.
For the small values of $p_{taes}$ likely to be encountered, this approximates exponential growth.

An approximation of the posterior distribution is computed using INLA (the Integrated Nested Laplace Approximation), see \cref{E-transmission:sec:INLA} for details.
I fitted the model separately and independently for each region, implicitly adding all interactions between region and the other variables.
This step assumes that the estimates for each region can be considered independent, which is reasonable because the regions are large with many tests conducted within each.

The final step is poststratifying the incidence proportion to the CIS target population and subpopulations of interest.
To start, I draw a sample of 500 prevalence time series in each stratum from the approximate posterior distribution.
For each posterior draw and day, the incidence proportion in the (sub)population is the weighted average of the incidence proportion in each stratum, where the weights are the number of individuals in that stratum.
The population numbers for each stratum are provided by the Office for National Statistics.

@@

Analyses requiring more fine-grained stratification (such as that in \cref{biology-data:sec:MRP}) or linking data from the same individual over time (such as that in \cref{E-perf-test,E-imperf-test}) must be performed within the SRS.
The SRS contains five 2.40\,GHz dual-core Intel Xeon E5-4650 processors and 164\,GB RAM.
These computational resources are shared between all users of the CIS SRS project and due consideration must be given to other users.
Processes which run for multiple days are liable to be cancelled.

Therefore, computational approaches requiring high-performance computing, graphical processing units, or long runtimes are not feasible within the SRS.
This computational limitation will feature in several places in this thesis.
